As mentioned above our prefetcher is a stride prefetcher. It tries to detect a 
pattern in the access history, and use this pattern to predict future 
references. Patterns could be detected on a global or a local (per PC) basis.
After reviewing statistics gathered from SPEC CPU2000 benchmarks we decided to 
focus on local patterns. This was because the global access stream had little 
additional unique information that could not already be extracted from the 
local streams without using a lot of additional resources. In most benchmarks 
the memory accesses are done by a few instructions that operate in a very 
predictable fashion. The prefetcher is very similar to a DCPT, but we store just the 
addresses instead of the delta values, and nothing else.

\subsection{Table}

\begin{figure}
	\begin{tikzpicture}
		\draw [->] (1,7.5) -- (2, 7.5);
		\draw (1.5, 7.8) node {PC};
		\foreach \x in {6,...,8}
			\draw (2, \x) rectangle (3, \x + 1);

		\draw [dotted] (2, 5.75) -- (2, 5.25);
		\draw [dotted] (3, 5.75) -- (3, 5.25);
		\draw (2, 5) -- (2, 5.25);
		\draw (2, 6) -- (2, 5.75);
		\draw (3, 5) -- (3, 5.25);
		\draw (3, 6) -- (3, 5.75);
		\draw (2, 4) rectangle  (3, 5);

		\foreach \x in {0,2}
			\draw (\x * 1.5 + 4, 6.75) rectangle  (\x * 1.5 + 5.5, 8.25);

		\draw [dashed] (3, 8) -- (4, 8.25);
		\draw [dashed] (3, 7) -- (4, 6.75);

		\draw [dotted] (5.5, 6.75) rectangle (7, 8.25);
		\draw (5.5, 6.75) -- (5.75, 6.75);
		\draw (6.75, 6.75) -- (7, 6.75);
		\draw (5.5, 8.25) -- (5.75, 8.25);
		\draw (6.75, 8.25) -- (7, 8.25);

		\draw (4.75, 7.5) node {$a_0$};
		\draw (7.75, 7.5) node {$a_{2n}$};
	\end{tikzpicture}
	\caption{Access History Table}
	\label{fig:table}
\end{figure}

To keep track of the previous memory accesses of a program counter we use a
table. Each table entry contains a list of the last memory accesses and misses
by the program counter that maps to this entry. On a hit or a miss the address
of the access is added at the front of the list corresponding to the program
counter of the instruction that made the access while the rest if shifted one
place to the right. In hardware this would be implemented with a circular buffer. A
graphical representation of the table can be seen in Fig.~\ref{fig:table}. The
exact number of accesses to store in each entry can be tuned. More addresses
per entry allows larger patterns to be detected, but reduces the number of
entries using the same amount of memory. Larger patterns might allow you to
catch more, while more entries reduces the chance of collision.

To save space we store only the lower $k$ bits in every address. Because the
simulated computer uses 64 byte cache lines, the lower 6 bits of addresses in
the cache will always be 0, and can safely be ignored. With $k=16$ we store
only 10 bits and can correctly identify patterns as long as the biggest stride
is less than $2^{16}$.

Because the table size is much smaller than the number of possible program
counters, several instructions could use the same table entry. If this happened
in a real program it could lead to less than optimal prefetching. With a
sufficiently large table the chance of two instructions clashing would be
negligible. In addition the table entry is found by taking the program counter
modulo the table size. This ensures that instructions that is less than table
size apart will map to different table entries, meaning all instructions in a
small enough loop will map to different table entries.

The reason for these limitations on the table, other than the obvious benefit
of using less space, was that the course staff had set down a 8KB limit on
memory usage.

\subsection{Pattern Matching}

When the table begins to fill we are able to find patterns in the memory accesses.
To find these patterns we calculate relative and absolute strides.
The relative stride is the difference between two consecutive
addresses. Absolute strides are calculated by choosing a pattern size $n$,
and then computing the difference between the address at position
$n$ to every address before $n$, and then doing the same for the address
at position $2n$.

That is,
\[
	s_i = \begin{cases}
		a_i - a_n,    & \text{if } i  < n \\%\in \{0 \dots n -1\}\\
		a_i - a_{2n}, & \text{if } i \geq n % \{n \dots 2n - 1\}
	\end{cases}
\]
where $s_i$~is the $i$th absolute stride, and $a_i$~is the address at position
$i$ in the table. Note that the newest address is at index 0 in the table. A
perfect match is found if
\begin{equation}
\label{eq:match}
\forall i \in \{0 \dots n - 1\}, \quad s_i = s_{i + n}
\end{equation}

It is also possible to find a partial match. We then simply relax the
requirement in equation \eqref{eq:match} that all strides have to match, to
that at least some percentage has to match.

An example of a perfect match can be seen in Table~\ref{table:pattern}.

Partial matching is the reason for using absolute strides instead of the more
intuitive relative stride. If we have a partial match, for example in a pattern
$a, b, x, d, a, b, y, d$ of length four with relative strides we can prefetch
with $a$ and $b$, but when it comes to $d$ we can not use it since it is
relative to the address after $x$ and $y$ which is not the same. However, if we
used absolute strides,~$d$~would be relative to the same address as $a$ and $b$
and could easily be used to prefetch.

\begin{table}[htb]
	\caption{Example of pattern with n=4}
	\label{table:pattern}
	\centering
	\begin{tabular}{c|c|c|c|c}
		\bfseries Position &
		\bfseries Address &
		\bfseries Stride &
		\bfseries Stride &
		\bfseries Stride \\
		& &
		\bfseries (Relative) &
		\bfseries (Absolute) &
		\bfseries Index \\
		\hline
		0 & 896   & & \\
		  &	& 64 & 448 & 0 \\
		1 & 832  & & \\
		  & & 64 & 384 & 1 \\
		2 & 768 & & \\
		  & & 128 & 320 & 2 \\
		3 & 640 & & \\
		  & & 192 & 192 & 3 \\
		4 & 448 & & \\
		  &	& 64 & 448 & 4 \\
		5 & 384 & & \\
		  &	& 64 & 384 & 5 \\
		6 & 320 & & \\
		  &	& 128 & 320 & 6 \\
		7 & 192 & & \\
		  &	& 192 & 192 & 7 \\
		8 & 0 & & \\
	\end{tabular}
\end{table}

\subsection{Most Common Stride}

In addition to the pattern matching above, we also implemented most common stride
prefetching. This was added as a fallback measure if the pattern matching code
was unable to find a pattern.
It looks for the most common relative stride, and if the number
of this stride is over a threshold
it will use it to prefetch a
constant number of cache lines.
In Table~\ref{table:pattern} the most common stride is 64.

\subsection{Algorithm}\label{sec:algorithm}

Pseudocode for the algorithm used can be found in Algorithm~\ref{alg:stride}.
The algorithm first tries to find a pattern of size $n$, then $n-1$ and so
on until it tries to find a pattern of size $1$. If no such patterns are
found it goes on to try finding a most common stride.

As we can see there are a few constants in the algorithm that can be
tuned. These are

~
\begin{itemize}
	\item \textbf{ THRESHOLD\subscript{stride}} \\ Determines when there is a match.
	For perfect matching this should be equal to the pattern size \\
	\item \textbf{ AGGR\subscript{stride}} \\ The aggressiveness determines how many
	cache lines we should prefetch. This is also called prefetch degree \\
	\item \textbf{ THRESHOLD\subscript{mcs}} \\ Sets a lower limit on how many
	there must be of the most commons stride before starting to prefetch \\
	\item \textbf{ AGGR\subscript{mcs}} \\  Same as the stride equivalent \\
\end{itemize}

It is important to note that these ``constants'' are actually C macros in our
source code, so they could depend on any variables in scope on the location
they are used.
Specifically \textbf{THRESHOLD\subscript{stride}} and
\textbf{AGGR\subscript{stride}} could depend on the pattern size.

\algnotext{EndIf}
\algnotext{EndFor}
\algnotext{EndFunction}

\begin{algorithm}
	\caption{The prefetching algorithm}
	\label{alg:stride}
	\begin{algorithmic}
		\Function{prefetch}{$a_0$, $a_1$, \dots, $a_{2n}$}
		\If{$\neg$ \textsc{stride\_prefetch}($a_0$, $a_1$, \dots, $a_{2n}$)}
			\State \textsc{mcs\_prefetch}($a_0$, $a_1$, \dots, $a_{2n}$)
		\EndIf
		\EndFunction
		\\
		\Function{stride\_prefetching}{$a_0$, $a_1$, \dots, $a_{2n}$}
			\For{$i = n \to 1$}
				\State $hits \gets 0$
				\For{$j = 0 \to i - 1$}
					\State $hit_j \gets$ \textnormal{false}
					\State $s_j \gets a_j - a_i$
					\State $s_{i + j} \gets a_{i + j} - a_{2i}$
				\EndFor
				\For{$j = 0 \to i - 1$}
					\If{$s_j \equiv s_{i + j}$}
						\State $hits \gets hits + 1$
						\State $hit_j \gets$ \textnormal{true}
					\EndIf
				\EndFor
				\If{$hits \geq$ THRESHOLD\subscript{stride}}
					\For{$j = 0 \to$ AGGR\subscript{stride}}
						\State $k \gets i - 1 - mod(i,j)$
						\If{$hit_k$}
							\State \textsc{prefetch}($a_0 + s_k + s_0 \cdot (\lfloor \frac{j}{i} \rfloor)$)
						\EndIf
					\EndFor
					\State {\bfseries return} \textnormal{true}
				\EndIf
			\EndFor

			\Return \textnormal{false}
		\EndFunction
		\\
		\Function{mcs\_prefetching}{$a_0$, $a_1$, \dots, $a_{2n}$}
			\State $mcs \gets s_0$
			\State $count \gets 0$

			\For{$i = 0 \to 2n$}
				\State $s_i \gets a_i - a_{i + 1}$
			\EndFor

			\For{$i = 0 \to 2n$}
				\State $c \gets 0$
				\State $m \gets s_i$
				\For{$j = i \to 2n$}
					\If{$s_i \equiv m$} \State $c \gets c + 1$ \EndIf
				\EndFor

				\If{$c > count$}
					\State $count \gets c$
					\State $mcs \gets m$
				\EndIf
			\EndFor
			\\
			\If{$count \geq$ THRESHOLD\subscript{mcs}}
				\For{$i = 1 \to$ AGGR\subscript{mcs}}
				\State \textsc{prefetch}($a_0 + mcs \cdot i$)
				\EndFor
			\EndIf
		\EndFunction
	\end{algorithmic}
\end{algorithm}
